---
title: "Cannibis Optimization Examples"
author: "Jerod Junkins"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, comment = "")
```

The Cloud Cannabis Company (Cloud) operates 4 dispensaries within Michigan, with a 5th opening soon in Detroit. Like most business in this space, they are faced with intense competition and an ever-growing consumer base.

## Stocking Shelves

Across the 4 current locations, Cloud offers product categories such as flower, pre-rolls, concentrates, edibles, and hardware. 

Let's imagine that each of the following investment strategies has some type of yearly ROI percent, assessed risk factor, and terms of investment in years:

For each of these product types, ROI, risk assessment, and cost are provided:

```{r}
data.frame(Product = c("Concentrates", "Hardware", "Pre-rolls", 
                        "Edibles", "Flower"), 
           ROI = c(12, 10, 15, 25, 0),
           Risk = c(2, 1, 3, 4, 0), 
           Cost = c(4, 8, 2, 10, 0)) |>
  flextable::flextable(cwidth = 2.5, cheight = .15)
```

Cloud wants to maximize the return, but overall risk should not exceed 2.5. Cloud does not want to exceed a cost value of 6 and they need at least 15% of the product to be in flower. Of Cloud's total product, what proportion should be given to each product type?

```{r}
library(linprog)

objectiveFunction <- c(12,10,15,25,0)

constraintValues <- c(2.5, 6, .15, 1)
constraintMatrix <- rbind(c(2,1,3,4,0),
                          c(4,8,2,10,0),
                          c(0,0,0,0,1),
                          c(1,1,1,1,1))

res <- solveLP(cvec = objectiveFunction, 
               bvec = constraintValues, 
               Amat = constraintMatrix, 
               maximum = TRUE, 
               const.dir = c("<=", "<=", ">=", "<="))

res$opt

res$solution
```

## Moving Product

Currently, Cloud receives products from the following companies: Better Provisioning (Kalamazoo), Harbor Farmz (Kalamazoo), Hazy Farms (Detroit), and Rare Michigan Genetics (Kalamazoo). Due to different locations, each company has slightly different flat shipping rates to each of the 4 current locations:

```{r, results='asis'}
data.frame(company = c(rep("Better Provisioning", 4), 
                       rep("Harbor Farmz", 4), 
                       rep("Hazy Farms", 4), 
                       rep("Rare Michigan Genetics", 4)), 
           store = rep(c("Ann Arbor", "Muskegon", "Utica", "Traverse City"), 4), 
           cost = c(10, 8, 12, 20, 
                    10, 8, 12, 20, 
                    5, 10, 5, 25,
                    10, 8, 12, 20)) |>
  flextable::flextable(cwidth = 2.5, cheight = .15)
```

In addition to the shipping costs to each location, each store has needed flower product and each grower only has so much product. Those values are listed below:

```{r}
data.frame(company = c("Better Provisioning", "Harbor Farmz",  
                       "Hazy Farms", "Rare Michigan Genetics"),  
           supply = c(700, 300, 550, 420), 
           store = c("Ann Arbor", "Muskegon", "Utica", "Traverse City"), 
           demand = c(650, 200, 450, 250)) |>
  flextable::flextable(cwidth = 2.5, cheight = .15)
```
Set up the various numbers
```{r}
cMat <- c(10, 8, 12, 20, 10, 8, 12, 20, 5, 10, 5, 25, 10, 8, 12, 20)

b <- c(700, 300, 550, 420, 650, 200, 450, 250)

A <- rbind(c(1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0),
  c(0, 0, 0, 0, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0),
  c(0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 0, 0, 0, 0),
  c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1),
  c(1, 0, 0, 0, 1, 0, 0, 0, 1, 0, 0, 0, 1, 0, 0, 0),
  c(0, 1, 0, 0, 0, 1, 0, 0, 0, 1, 0, 0, 0, 1, 0, 0),
  c(0, 0, 1, 0, 0, 0, 1, 0, 0, 0, 1, 0, 0, 0, 1, 0),
  c(0, 0, 0, 1, 0, 0, 0, 1, 0, 0, 0, 1, 0, 0, 0, 1))
```

Solve the product moving problem
```{r}
library(ROI)
library(ROI.plugin.glpk)

constraints <- L_constraint(A, c(rep("<=", 4), rep(">=", 4)), b)

model <- OP(objective = cMat,
            constraints = constraints,
            types = rep.int("I", length(cMat)),
            maximum = FALSE)

result <- ROI::ROI_solve(model, "glpk", verbose = TRUE)
```
Get the final values
```{r}
transportation_solution <- solution(result)

names(transportation_solution) <- c("c1_s1", "c1_s2", "c1_s3", "c1_s4",
                                    "c2_s1", "c2_s2", "c2_s3", "c2_s4",
                                    "c3_s1", "c3_s2", "c3_s3", "c3_s4",
                                    "c4_s1", "c4_s2", "c4_s3", "c4_s4")

result$objval
transportation_solution
```
From Better Provisioning, send 550 units to Ann Arbor and 150 units to Muskegon

From Harbor Farmz, send 50 units to Muskegon and 250 units to Traverse City

From Hazy Farms, send 100 units to Ann Arbor and 450 units to Utica

Nothing will be sent from Rare Michigan Genetics

The total cost will be $14,850

## From Sales to Scales

As with most dispensaries, Cloud is contemplating a leap into the production space. Unfortunately, it is not without a fair amount of risk. The risk is not about money (although the automation involved is costly), but about time. Generally, the following products get created from a plant: flower and concentrates. 

From start to finish, the process to produce flower is as follows:

1. Flowering stage (min: 8 weeks; max: 11 weeks)

2. Drying stage (min: 7 days; max: 10 days)

3. Trimming stage (mu = 1 hour; sigma = 15 minutes)

4. Curing stage (min: 1 month; max: 4 months)

From start to finish, the process to produce concentrates is as follows:

1. Flowering stage (min: 8 weeks; max: 11 weeks)

2. Drying stage (min: 7 days; max: 10 days)

3. Freeze stage (20 minutes)

4. Extraction stage (mu: 15 minutes; sigma: 2 minutes)

5. Filtration stage (mu: 1 hour; sigma: 8 minutes)

6. Solvent evaporation stage (1 hour)

7. Decarbonization stage (mu: 2 days; sigma: 8 hours)

8. Distillation stage (mu: 18 hours; sigma 2 hours)

About 45% of a given grow goes toward flower, while the remaining 55% goes towards concentrates. How long should Cloud anticipate a first grow taking?

```{r}
library(DiagrammeR)
library(simmer)
library(simmer.plot)

sim1000runs <- replicate(1000, expr = {

product <- trajectory("parts") %>% 
  set_attribute("start_time", function() {now(production)}) %>% 
  seize("flowering stage") %>% 
  timeout(function(){runif(n = 1, min = 56, max = 77)}) %>% 
  release("flowering stage") %>% 
  seize("drying stage") %>% 
  timeout(function(){runif(n = 1, min = 7, max = 10)}) %>% 
  release("drying stage") %>% 
  
branch(function() {sample(1:2, 1, prob = c(.45, .55))}, 
         continue = c(TRUE, TRUE), 
         trajectory() %>% 
           seize("trimming stage") %>% 
           timeout(function() {rnorm(n = 1, mean = 1/24, sd = 1/96)}) %>% 
           release("trimming stage") %>% 
           seize("curing stage") %>% 
           timeout(function(){runif(n = 1, min = 30.437, max = 121.748)}) %>% 
           release("curing stage"),
        trajectory() %>% 
           seize("freeze stage") %>% 
           timeout(1/72) %>%  
           release("freeze stage") %>% 
           seize("extraction stage") %>% 
           timeout(function() {rnorm(n = 1, mean = 1/96, sd = 1/720)}) %>% 
           release("extraction stage") %>% 
           seize("filtration stage") %>% 
           timeout(function() {rnorm(n = 1, mean = 1/24, sd = 1/180)}) %>% 
           release("filtration stage") %>% 
           seize("solvent evaporation stage") %>% 
           timeout(1/24) %>% 
           release("solvent evaporation stage") %>% 
           seize("decarbonization stage") %>% 
           timeout(function() {rnorm(n = 1, mean = 2, sd = 1/3)}) %>% 
           release("decarbonization stage") %>% 
           seize("distillation stage") %>% 
           timeout(function() {rnorm(n = 1, mean = 3/4, sd = 1/12)}) %>% 
           release("distillation stage"))

production <- simmer("production") %>% 
  add_resource("flowering stage") %>%
  add_resource("drying stage") %>% 
  add_resource("trimming stage") %>% 
  add_resource("curing stage") %>% 
  add_resource("freeze stage") %>% 
  add_resource("extraction stage") %>% 
  add_resource("filtration stage") %>% 
  add_resource("solvent evaporation stage") %>% 
  add_resource("decarbonization stage") %>% 
  add_resource("distillation stage") %>% 
  add_generator("grow", product, at(0))

simmer::run(production, until = 1000)

result <- get_mon_arrivals(production)

}, simplify = FALSE)

simulations <- do.call(rbind, sim1000runs)
simulations
mean(simulations$end_time)

grow_time <- simulations$end_time
hist(grow_time,
     main="Average Grow Time",
     xlab="Days Elapsed",
     ylab="Count",
     col="darkgreen")

```


## Return On Investment

The cost to start plant production is high. The following table offers a break down of some yearly cost range estimates for a 10,000 square foot greenhouse:

```{r}
data.frame(item = c("Rent", "Lighting", "Climate Control", 
                    "Irrigation", "Consumables"),  
           low = c(100000, 5000, 10000, 2500, 25000), 
           mid = c(150000, 10000, 15000, 3000, 50000), 
           high = c(300000, 20000, 17000, 5000, 75000)) |>
  flextable::flextable(cwidth = 2.5, cheight = .15)
```

A high-functioning greenhouse can churn out an average of 1,400 pounds of dried product per year. However, a great number of things can happen: power failures, pests, poor yield. While the average might be 1,400 pounds, there is some variance around that (roughly 300 pounds as a standard deviation). If 1,400 pounds will return a gross profit of 1,000,000 million dollars, how much would Cloud likely make over the next 5 years?  

Libraries
```{r}
library('dplyr')
library('ggplot2')
library('triangle')
```

Loop for cost
```{r}
reps_cost <- replicate(n = 1000, expr = {
n_years <- 5
total_cost <- 0

for (i in 1:n_years){

rent_draws <- rtriangle(n = 1, a = 100000, b = 300000, c = 150000)
lighting_draws <- rtriangle(n = 1, a = 5000, b = 20000, c = 10000)
climate_draws <- rtriangle(n = 1, a = 10000, b = 17000, c = 15000)
irrigation_draws <- rtriangle(n = 1, a = 2500, b = 5000, c = 3000)
consumables_draws <- rtriangle(n = 1, a = 25000, b = 75000, c = 50000)

total_cost = total_cost + rent_draws + lighting_draws + climate_draws + irrigation_draws + consumables_draws

i = i + 1}
  results <- data.frame(total_costs = total_cost,
                        years = n_years)
  
  results
}, simplify = TRUE)

hist(unlist(reps_cost["total_costs",])
     , xlab = "Total Cost")
```

Loop for revenue
```{r}
reps_revenue <- replicate(n = 1000, expr = {
n_years = 5
total_pounds = 0 

for (i in 1:n_years){
  yield_draws <- rnorm(1, 1400, 300)
  
  total_pounds <- total_pounds + yield_draws
i = i + 1}

total_revenue = total_pounds* (1000000/1400)

  results <- data.frame(years = n_years,
                        total_revenues = total_revenue)
  
  results
}, simplify = TRUE)

hist(unlist(reps_revenue["total_revenues",])
     , xlab = "Total Revenue")
```

